---
title: "biosurvey: Tools for Biological Survey Planning"
author: "Claudia Nunez-Penichet, Marlon E. Cobos, A. Townsend Peterson, Narayani Barve, Vijay Barve, Tomer Gueta"
output: 
  github_document:
    toc: yes
    toc_depth: 3
always_allow_html: true
---

```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

<br>

**This repository is for the project "Biological Survey Planning Considering Hutchinson’s Duality" developed during the program GSoC 2020.**

<br>

## Project description

Student: *Claudia Nuñez-Penichet*

GSoC Mentors: *Narayani Barve, Vijay Barve, Tomer Gueta*

Motivation:

Given the increasing intensity of threats to biodiversity in the world, one of the challenges in biodiversity conservation is to complete inventories of existing species at distinct scales. Species distributions depend on the relationships between accessible areas, environmental conditions, and biotic interactions. As planning a survey system only aims to register species in a region, biodiversity interaction can be overlooked in this case. However, the relationship between environmental conditions and the geographic configuration of an area is of crucial importance when trying to identify key sites for biodiversity surveys.  Among the diverse packages in R for selecting survey sites,  such considerations are not implemented and are limited to a random selection of sampling sites or analyses that allow detecting potential sampling sites based on the environmental similarity between sampled and unsampled areas. Given the need for more solutions, the  **biosurvey** package aimed for considering the relationship between environmental and geographic conditions in a region when designing survey systems that allow sampling of most of its biodiversity. 

<br>

## Package description

The biosurvey R package implements multiple tools to allow users to select sampling sites increasing efficiency of biodiversity survey systems by considering the relationship of environmental and geographic conditions in a region.

<br>

## Installing the package

biosurvey is in a GitHub repository and can be installed and/or loaded using the code below (make sure to have Internet connection). If you have any problem during installation, restart R session, close other RStudio essions you may have open, and try again. If during the installation you are asked to update packages, do so if you don't need a specific version of one or more of the packages to be installed. If any of the packages gives an error when updating, please install it alone using install.packages(), then try re-installing biosurvey again.

```{r installation, eval=FALSE}
# Installing and loading packages
if(!require(remotes)){
  install.packages("remotes")
}
if(!require(biosurvey)){
  remotes::install_github("claununez/biosurvey")
  library(biosurvey)
}
```
